### 2013-09-15 doc/chapters/overview.tex
add manual tex file with the rough structure layout

\section{Overview}

Explain:
\begin{itemize}
\item What is FlexibleSUSY?
\item Features, design goals
\end{itemize}

\subsection{Requirements}

\begin{itemize}
\item Mathematica 7-9
\item SARAH version 4
\item a C++11 compatible compiler (g++ 4.7.2 or higher, clang++ 3.1 or
  higher)
\item a fortran compiler (gfortran etc.)
\item Eigen library version 3
\item Boost library
\item GNU scientific library
\item lapack (needed for the Lattice algorithm only)
\end{itemize}