### 2013-09-15 doc/chapters/quick_start.tex

\section{Quick start}
Create a MSSM spectrum generator
  \begin{lstlisting}[language=bash]
$ ./createmodel --models=MSSM
$ ./configure --with-models=MSSM
$ make
  \end{lstlisting}
Run the spectrum generator using the default parameter point
  \begin{lstlisting}[language=bash]
$ ./models/MSSM/run_MSSM.x
  \end{lstlisting}
Run the spectrum generator using a SLHA input file
  \begin{lstlisting}[language=bash]
$ ./models/MSSM/run_MSSM.x --slha-input-file=templates/MSSM/LesHouches.in.MSSM --slha-output-file=LesHouches.out.MSSM
  \end{lstlisting}