### 2013-09-15 doc/chapters/usage.tex

\section{Usage}
\subsection{Files and directories}

Explain:
\begin{itemize}
\item Soucre code directories and files
\item Model directories and files
\end{itemize}

\subsection{Basic commands}

Explain:
\begin{itemize}
\item ./createmodel
\item ./configure
\item make
\item Mathematica interface (see models/<model>/start.m)
\end{itemize}

\subsection{Model files}

Explain how to write a FlexibleSUSY model file (starting from a SARAH
model file).

\subsection{Makefile modules}

Explain how to write a custom makefile module (module.mk) for a
FlexibleSUSY model.