### 2013-09-16 doc/chapters/usage.tex
add overview for the FlexibleSUSY model file variables

\section{Usage}
\subsection{Files and directories}

Explain:
\begin{itemize}
\item Soucre code directories and files
\item Model directories and files
\end{itemize}

\subsection{Basic commands}

Explain:
\begin{itemize}
\item ./createmodel
\item ./configure
\item make
\item Mathematica interface (see models/<model>/start.m)
\end{itemize}

\subsection{Model files}

Explain how to write a FlexibleSUSY model file (starting from a SARAH
model file).

A FlexibleSUSY model file is a Mathematica file, where the following
pieces are defined:

\paragraph{General model information}

\subparagraph{FSModelName}
Name of the FlexibleSUSY model.
Example (MSSM):
\begin{lstlisting}[language=Mathematica]
FSModelName = "MSSM";
\end{lstlisting}

\subparagraph{OnlyLowEnergyFlexibleSUSY} If set to True, creates a
spectrum generator without high-scale constraint, i.e.\ only a
low-scale and susy-scale constraint will be created (default: False).
In this case all model parameters, except for gauge and Yukawa
couplings are input at the susy scale.  This option is similar to
\texttt{OnlyLowEnergySPheno} in SARAH/SPheno.
\\
Example:
\begin{lstlisting}[language=Mathematica]
OnlyLowEnergyFlexibleSUSY = False;  (* default *)
\end{lstlisting}

\paragraph{Input and output parameters}

\subparagraph{MINPAR} This list of two-component lists defines model
input parameters and their SLHA keys.  These parameters will be read
from the MINPAR block in the SLHA file.
\\
Example (MSSM):
\begin{lstlisting}[language=Mathematica]
MINPAR = { {1, m0},
           {2, m12},
           {3, TanBeta},
           {4, Sign[\[Mu]]},
           {5, Azero}
         };
\end{lstlisting}

\subparagraph{EXPAR} This list of two-component lists defines further
model input parameters and their SLHA keys.  These parameters will be
read from the EXTPAR block in the SLHA file.
\\
Example (E6SSM):
\begin{lstlisting}[language=Mathematica]
EXTPAR = { {61, LambdaInput},
           {62, KappaInput},
           {63, muPrimeInput},
           {64, BmuPrimeInput},
           {65, vSInput}
         };
\end{lstlisting}

\subparagraph{DefaultParameterPoint} This list of two-component lists
defines default values for the model input parameters.
\\
Example (MSSM):
\begin{lstlisting}[language=Mathematica]
DefaultParameterPoint = {
    {m0, 125},
    {m12, 500},
    {TanBeta, 10},
    {Sign[\[Mu]], 1},
    {Azero, 0}
};
\end{lstlisting}

\paragraph{Constraints}
Currently three constraints are supported: low-scale, susy-scale and
high-scale constraints.  In FlexibleSUSY they are named as
\texttt{LowScale}, \texttt{SUSYScale} and \texttt{HighScale}.  For
each constraint there is a scale definition (named after the
constraint), an initial guess for the scale (concatenation of the
constraint name and \texttt{FirstGuess}) and a list settings to be
applied at the constraint (concatenation of the constraint name and
\texttt{Input}).
\\
Example (MSSM):
\begin{lstlisting}[language=Mathematica]
(* susy-scale constraint *)
SUSYScale = Sqrt[M[Su[1]]*M[Su[6]]];        (* scale definition *)
SUSYScaleFirstGuess = Sqrt[m0^2 + 4 m12^2]; (* first scale guess *)
SUSYScaleInput = {};                        (* nothing is set here *)

(* high-scale constraint *)
HighScale = g1 == g2;                       (* scale definition *)
HighScaleFirstGuess = 1.0 10^14;            (* first scale guess *)
HighScaleInput = {
   {T[Ye], Azero*Ye},
   {T[Yd], Azero*Yd},
   {T[Yu], Azero*Yu},
   {mHd2, m0^2},
   {mHu2, m0^2},
   {mq2, UNITMATRIX[3] m0^2},
   {ml2, UNITMATRIX[3] m0^2},
   {md2, UNITMATRIX[3] m0^2},
   {mu2, UNITMATRIX[3] m0^2},
   {me2, UNITMATRIX[3] m0^2},
   {MassB, m12},
   {MassWB,m12},
   {MassG, m12}
};

LowScale = SM[MZ];                          (* scale definition *)
LowScaleFirstGuess = SM[MZ];                (* first scale guess *)
LowScaleInput = {
   {vd, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] Cos[ArcTan[TanBeta]]},
   {vu, 2 MZDRbar / Sqrt[GUTNormalization[g1]^2 g1^2 + g2^2] Sin[ArcTan[TanBeta]]}
};
\end{lstlisting}

\paragraph{Initial parameter guesses} At the \texttt{LowScale} and
\texttt{HighScale} it is recommended to make an initial guess for the
model parameters.  This can be done via
\begin{lstlisting}[language=Mathematica]
InitialGuessAtLowScale = {
   {vd, SM[vev] Cos[ArcTan[TanBeta]]},
   {vu, SM[vev] Sin[ArcTan[TanBeta]]}
};

InitialGuessAtHighScale = {
   {\[Mu]   , 1.0},
   {B[\[Mu]], 0.0}
};
\end{lstlisting}

\subsection{Makefile modules}

Explain how to write a custom makefile module (module.mk) for a
FlexibleSUSY model.